---
title: "Rebeca: intermediate R/eurostat user"
subtitle: "Release a simple, FAIR dataset obtained with the eurostat package."
format: html
editor: visual
---

## Eurostat user

This is a usability and requirement scenario for [Making Datasets Truly Interoperable in R](https://music.dataobservatory.eu/documents/open_music_europe/dataset-development/dataset-working-paper.html) --- source:  [dataset-working-paper.qmd](https://github.com/dataobservatory-eu/dataset-development/stories/)

**Persona:** Rebeca is an intermediate R user who is familiar with the `tidyverse` and the `eurostat` package, and the basic concept of tidy data. She works with cultural statistics and wants to make her work FAIR and link to other scientists' work.

**Before:** Rebeca is able to download the _Cultural employment by sex_ Eurostat statistics with the _eurostat_ R pacakage:

```{r loadpackages, echo="FALSE", message=FALSE, warning=FALSE}
library(eurostat)
library(dplyr)
library(tidyr)
cult_emp_sex <- get_eurostat("cult_emp_sex")
str(cult_emp_sex)
```

... and create a derived dataset showing the annualised cultural employment growth rates of European countries from the original  [cult_emp_sex](https://ec.europa.eu/eurostat/databrowser/view/cult_emp_sex/default/table?lang=en) dataaset with _tidyverse_ functions.

```{r simplecalculus}

cult_emp_growth_sex <- cult_emp_sex %>% 
  filter (time %in% c(as.Date("2022-01-01"), as.Date("2013-01-01"))) %>%
  mutate ( time_var = paste0("Y", substr(as.character(time), 1,4))) %>%
  select (-time) %>%
  pivot_wider ( names_from = "time_var", values_from = "values") %>%
  mutate( annaul_growth = (Y2022/Y2013)^0.1 ) 

cult_emp_growth_sex 
```

Until now, she would save these results into a `.csv` file and upload it to her research website and Zenodo manually. The proaboutblem with this approach that users who download the `csv` file from her website do not have a clear idea what these variables stand for, or what is the provenance of the work. Such information is available on Zenodo repository, but users who download the `.csv` file may forget about it.

__After__: Rebeca can save the `cult_emp_growth_sex` dataset in R into an `.rds` file that contains all the DataCite or Dublin Core metadata that is required on Zenodo for publication. She can export this `cult_emp_growth_sex.rds` file into a CSV format that meets the W3C consortium's standard on publishing CSV with machine-readable JSON metadata. She can also serialize for long-term usability the `cult_emp_growth_sex` it into an RDF schema that contains all the semantic information to connect this dataset to other data that use the SDMX semantic standards, such as other datasets of Eurostat, the World Bank or OECD.

## Functional requirements

1. The _eurostat_ package retains the provenance metadata, i.e., the descriptive metadata and the semantics of the Eurostat original `cult_emp_sex`. _The dataset package is used by eurostat to record the metadata from the source_

2. The _eurostat_ package retains the valid range and the codelist of each variable, in this case, `sex`, `unit`, `geo`, `time`, and the measured values. _The dataset package is used by eurostat to record the codelists from the source_

3. Rebeca can add her own descriptive metadata, i.e., as the creator of the derived `cult_emp_growth_sex` dataset in a way that `cult_emp_growth` becomes a related item with the derivative work; Rebeca is recorded as the creator. 

4. The _eurostat_ package is added to the related items metadata as a software agent that was used in the creation of the `cult_emp_growth_sex` derived dataset.  _The dataset package has a function that adds related items metadata._

5. Rebeca can export the `cult_emp_growth_sex` dataset into an RDF resource which contains the data, its semantic description, its data structure, the provenance and related items data, and her desciptive metadata as required by FAIR.  _The dataset package has a release function that can be used._

6. Rebeca can export the `cult_emp_growth_sex` dataset into a standard CSV file with standard JSON metadata which contains the data, its semantic description, its data structure, the provenance and related items data, and her desciptive metadata as required by FAIR. _The dataset package has a release function that can be used._

7. Rebeca can release the package via the connector [zen4r](https://cran.r-project.org/web/packages/zen4R/index.html) to the Zenodo repository. _This is explained in a vignette article._




